---
title: "Homework 4"
author: "Olivia Bickford"
date: "4/20/2022"
output: html_document
---
April 21, 2022
Olivia Bickford 

This is homework 4
Working with the Collatz Sequence
Using more practice with conditional code 
  to loop together functions and create sequences and user specified code 

I also created plots and bar graphs and learned 
  more about how those functions worked and how I can create different graphs

In the first question I made a functions outlining 
  the specifics for collatz sequence such as is a number 
  is divided by 2 and equals 0 it returns that number, or 
  if not the number goes through 3*num +1 

Then using conditional code, I made a function that printed out the 
  full sequence for an n value used from 1-10000
  
Attempt at the bonus and sources referrenced are at the end!

```{r needed packages}
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("ggplot2")) install.packages("ggplot2")
library(ggplot2)
```

These are the packages needed for the following code.

## Directions for Setting Up Collatz Sequence

```{r directions }
directions <- function(num) { ##basic collatz sequence, outputs next number in the sequence
  if(num == 1) { ## and sets up the basic function for the collatz sequence
    return(NULL)
  } 
  
  if(num %% 2 == 0) {
    return(num/2)
  } 
  else{
    return(3*num+1)
  }
}
```


## Collatz Loop Squences for a given value n

```{r collatz sequence}
collatzloop_sequences <- function(n) { ## Prints out the collatz sequence for a given n value 
  stated_problem <- n
  while (n != 1) {
    n <- directions(n=n)
    stated_problem <- c(stated_problem, n)
  }
  return(stated_problem)
}
collatzloop_sequences(6)  ## put in 6 for an example 
```


## Histogram for Collatz Sequence lengths 1-10000

```{r histogram for collatz lengths}
collatzloop <- c()
for (i in 1:10000) {
  collatzloop <- c(collatzloop, length(collatzloop_sequences(i)))
}
hist(collatzloop)
```

The histogram for 10,000 sample lengths and the frequency is above. This histogram is right skewed and bimodal. This means that most of the sample lengths are in the lower numbers such as 60,70,80 as compared to higher numbers such as 250. From the summary statistics below you can see that the mean sample length is 85.97. The highest length is 262 and the lowest is 1 which is expected.You have more sample lengths that are in the lower numbers which pulls the distribution and skews it to the right. 


## Summary Statistics 

```{r summary}
summary(collatzloop)
```


## Bonus Question

```{r bonus}
collatzloop_2_bonus <- c() ## Save it in a vector
for (i in 1:10000) {
  collatzloop_2_bonus <- c(collatzloop_2_bonus, (collatzloop_sequences(i)))
}
collatzloop_bonus <- tabulate(collatzloop_2_bonus, nbins = 100) ##tables this and puts into bins to count the sequence only integers 1-100 and show how many times that number occurs in the collatz sequence from a sample of 1-10000
collatzloop_bonus 
```

1 appears 10,000 times since it is the end of every sequence.

## BarGraph

```{r bargraph bonus}
integers <- (1:100)
frequency_of_integers <- collatzloop_bonus

collatzloop_bonus_df <- as.data.frame(collatzloop_bonus) ##made this a data frame for integers 1-100 and put collatzloop_bonus into a data frame 

ggplot(data = collatzloop_bonus_df, aes(x= integers, y = frequency_of_integers )) + 
  geom_col(fill= "lightblue", colour="black") ##build bar plot with the two variables above
```


## The Number between 1 and 100 that appears the least amount of times

```{r bonus least amount}
minimum_bonus <- which(collatzloop_bonus == min(collatzloop_bonus)) ## find minimum frequencies of the data frame and the specific integers that go along with the frequencies
minimum_bonus
```

These numbers all have the same frequency and they all appear the least amount of times.

## The odd number (not counting one) that appears the most amount of times

```{r bonus odd num}
df_odd_integers <- data.frame(integers = 1:100, y =frequency_of_integers) ## put in data frame for odd numbers but put label of integers that correlate with the frequencies 

row_odd <- seq_len(nrow(df_odd_integers)) %% 2  ## select odd numbers        

odd_nums_df <- df_odd_integers[row_odd == 1, ]

arrange(odd_nums_df, desc(y)) %>%  ##arrange frequency of integers from highest to lowest
  slice(2) ##slice the second value
```

After 1, 5 appears the most amount of times. It appears 9400 times.

https://r-graphics.org/recipe-bar-graph-basic-bar
  Used this link to help guide in making the bar plot    
      -Learned about geom_col for colors and outlining
      -Attempted to add x and y variables to have labels for the axis'
      and have numeric labels to show frequency range and integers
      
https://stackoverflow.com/questions/9390749/return-index-of-the-smallest-value-in-a-vector
  Used this to help with finding minimum value in vector 
  
  